{\boldmath Pseudoscalar top-Higgs coupling: Exploration of $\mathrm{CP}$-odd observables to resolve the sign ambiguity }

### 2016-02-10 Paper_ttbarH_Draft.tex
Edits to abstract and first part of introduction

\author{Daniel Crane \vspace*{4mm}}
\email{\hlkk{dkcrane@mtu.edu}}
\tayloru

\author{Ethan Gegner}
\email{ethan_gegner@taylor.edu}
\tayloru

@@

\begin{abstract}
\vspace*{4mm} We present a collection of $\mathrm{CP}$-odd observables
for the process $\ppprocess$ that are linearly dependent on the scalar
($\kp$) and pseudoscalar ($\kpt$) top-Higgs coupling and hence
sensitive to the corresponding relative sign. The proposed observables
are based on triple products (TP) structures that we extract from the
expression of the differential cross section in terms of the spin
vectors of the top and antitop quarks. In order to explore other
possibilities, we progressively modify these TPs, firstly, by
combining them and then by replacing the spin vectors by the lepton
momenta or the $t$ and $\tbar$ momenta by their visible parts. We
obtain that the most promising observable can disentangle the
hypotheses $\kp=1,\kpt=\pm 1$ by more than $\sim 20\sigma$ for $10^5$
events that is within the expected order of magnitude of events at the
HL-LHC. In the case of observables that do not need the reconstruction
of the $t$ and $\tbar$ momenta, the power of discrimination is up to
$\sim 16\sigma$ for the same ammount of events. We also show that the
capability of the most promising observables for separating the
$\mathrm{CP}$-mixed hypotheses preveals even when a number of events
plausible within the short term LHC is considered.

@@

%We also show that even when the number of events is reduced by an order of magnitude the most promising observables are still useful for testing the couplings
\end{abstract}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec1}
After the discovery of a new boson $H$ by the ATLAS \cite{atlasH} and
CMS \cite{cmsH} collaborations, it has become of crucial importance
its characterization, namely the determination of its physical
properties with the highest possible precision. Among the features of
the new boson that are under investigation, the study of its couplings
to the fermions is of great relevance, allowing us to study the nature
of $H$ under the discrete symmetry $\mathrm{CP}$ and to clarify the
extent to which this new particle is consistent to the Higgs boson
predicted by the Standard Model (SM) of particle physics. In
particular, the test of the Higgs coupling to the heaviest fermion,
the top quark, is a priority since on the one hand, it rules the main
production mechanism through gluon fusion and contributes to the
important decay mode to two photons, and on the other hand, it is
involved in the scalar-field naturalness problem giving rise to the
leading dependence on the cut-off energy scale in the corrections to
the Higgs mass and it may play an important role in the EWSB
mechanism.\par

Given the fact that the main Higgs boson production
process, the gluon fusion, is dominated by a top quark loop and that
the diphoton and digluon decay channels are also mediated by a top
loop, these processes provide constraints on the scalar and
pseudoscalar $tH$ couplings, $\kp$ and $\kpt$
\cite{constraints1,constraints2, constraints3,constraints4}. However,
these constraints assume that there are no other sources contributing
to the corresponding effective couplings and, in the case of the
diphoton decay channel which also involves a $W$ boson loop, that the
coupling of the Higgs boson to the $W$ is standard. In this sense, the
constraints derived from measurements of Higgs rates are indirect
constraints. In this direction, electric dipole moments (EDM) can also
impose stringent indirect constraints on $\kpt$ by assuming that there
are no new physics (NP) particles contributing to the loops of the
relevant diagrams and that the electron-Higgs coupling is that
predicted by the SM \cite{constraints1,edm}. In order to probe
directly the $tH$ coupling, processes with smaller cross sections need
to be taken into account.  \par

@@

\newpage

@@

factor, we will concentrate in the following combination arising from

>>>>

\author{Daniel Crane \vspace*{4mm}}
\email{dkcrane@mtu.edu}
\tayloru

\author{Ethan Gegner}
\email{ethan\_gegner@taylor.edu}
\tayloru

@@

\begin{abstract}
\vspace*{4mm} We present a collection of $\mathrm{CP}$-odd observables
for the process $\ppprocess$ that are linearly dependent on the scalar
($\kp$) and pseudoscalar ($\kpt$) top-Higgs coupling and hence
sensitive to the corresponding relative sign. The proposed observables
are based on triple product (TP) structures that we extract
from the expression of the differential cross section in terms of the
spin vectors of the top and antitop quarks. In order to explore other
possibilities, we progressively modify these TPs, first
by combining them, and then by replacing the spin vectors by the
lepton momenta or the $t$ and $\tbar$ momenta by their visible parts.
Assuming an integrated luminosity that is consistent with that
  envisioned for the HL-LHC, we find that the most
promising observable can disentangle the hypotheses $\kp=1,\kpt=\pm 1$
by more than $\sim 20\sigma$.  In
the case of observables that do not need the reconstruction of the $t$
and $\tbar$ momenta, the power of discrimination is up to $\sim
16\sigma$ for the same number of events. We also show that the
capability of the most promising observables for separating the
$\mathrm{CP}$-mixed hypotheses preveals even when a number of events
plausible within the short term LHC is considered. {\bf (KK: reword last sentence a bit.)}

@@

%We also show that even when the number of events is reduced by an order of magnitude the most promising observables are still useful for testing the couplings
\end{abstract}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec1}
After the discovery of a new boson $H$ by the ATLAS \cite{atlasH} and
CMS \cite{cmsH} collaborations, it has become of crucial importance
 to determine its physical
 properties with the highest possible precision.
 The study of the new boson's couplings
 to fermions is of great relevance
 and will allow us to better understand this particle's $\mathrm{CP}$-transformation
   properties, as well as the
extent to which this particle is consistent with the Higgs boson
predicted by the Standard Model (SM) of particle physics.
It is particularly important to test the coupling of the putative ({\bf (KK: at some point
  should we just start calling this the Higgs?  How do we make that transition?)}
Higgs boson to
the top quark.  This coupling governs
the main Higgs boson production mechanism (which proceeds via gluon fusion)
and it contributes to the
important Higgs boson decay mode to two photons.
It is also
involved in the scalar-field naturalness problem -- giving rise to the
leading dependence on the cut-off energy scale in the corrections to
the Higgs mass -- and it may play an important role in the EWSB {\bf KK: define EWSB?}
mechanism.  {\bf (KK: I reworded a lot of this paragraph -- is it still
  correct, or did I accidentally alter the meaning?)}\par

Given that the main Higgs boson production
process is dominated by a top quark loop and that
the diphoton and digluon decay channels are also mediated by a top
loop, these processes provide constraints on the scalar and
pseudoscalar $tH$ couplings, $\kp$ and $\kpt$
\cite{constraints1,constraints2, constraints3,constraints4}.
     {\bf (KK: OK to drop ``gluon fusion'' here, since we already said it was
       the main production process above?  Alternatively, we could say it here
       and skip it above.)}
     However,
these constraints assume that there are no other sources contributing
to the corresponding effective couplings; furthermore, in the case of the
diphoton decay channel (which also involves a $W$ boson loop), it is also
assumed that the
coupling of the Higgs boson to the $W$ is standard. In this sense, the
constraints derived from measurements of Higgs rates are indirect
constraints. Electric dipole moments (EDM) can also
impose stringent indirect constraints on $\kpt$ by assuming that there
are no new physics (NP) particles contributing to the loops of the
relevant diagrams and that the electron-Higgs coupling is that
predicted by the SM \cite{constraints1,edm}. In order to probe
the $tH$ coupling directly, processes with smaller cross sections need
to be taken into account.  \par

@@

@@

factor, we will concentrate on the following combination arising from
